---
title: "Project 2"
author: "Vinh Van"
date: "4/16/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list=ls())
```

```{r}
grades <- read.csv(
  file = "C:\\Users\\Duyet\\Downloads\\School\\Spring 2022\\STA4164\\Project 2\\student-mat.csv"
)
```

##Comment out when knitting
```{r}
##grades
```

Summary Statistics and Histogram of G3
```{r}
summary(grades[c('G3')])
hist(grades[,'G3'],xlab='Final Grades',main='Histogram of G3')
```

Creating the scatter plots
```{r}
par(
  mfrow=c(2,3)
)

plot(
    x = grades$age,
    y = grades$G3,
    xlab = "Age",
    ylab = "Final Grades",
)
abline(
  reg = lsfit(
    x = grades$age,
    y = grades$G3
  )
)

plot(
    x = grades$failures,
    y = grades$G3,
    xlab = "# of Past Class Failures",
    ylab = "Final Grades",
)
abline(
  reg = lsfit(
    x = grades$failures,
    y = grades$G3
  )
)

plot(
    x = grades$absences,
    y = grades$G3,
    xlab = "Number of School Absences",
    ylab = "Final Grades",
)
abline(
  reg = lsfit(
    x = grades$absences,
    y = grades$G3
  )
)

plot(
    x = grades$G1,
    y = grades$G3,
    xlab = "First Semester Grades",
    ylab = "Final Grades",
)
abline(
  reg = lsfit(
    x = grades$G1,
    y = grades$G3
  )
)

plot(
    x = grades$G2,
    y = grades$G3,
    xlab = "Second Semester Grades",
    ylab = "Final Grades",
)
abline(
  reg = lsfit(
    x = grades$G2,
    y = grades$G3
  )
)
```



Creating a model of just the quantatative variables
```{r}
lm_initial <- lm(
  formula = G3 ~ G1 + G2 + age + failures + absences,
  data = grades,
)

lm_initialsquared <- lm(
  formula = G3^2 ~ G1 + G2 + age + failures + absences,
  data = grades,
)

summary(lm_initial)
summary(lm_initialsquared)

anova(lm_initial)
anova(lm_initialsquared)

```   

Correlation matrix
```{r}
mcor <- cor(grades[c('G3','G1','G2','age','failures','absences')])
mcor
```

Variance Inflation Factor
```{r}
library(car)
vif <- vif(lm_initial)
vif
```

Diagnostic Plots
```{r}
v_color <- viridis::viridis(
  n = nrow(
    x = grades
  )
)
grades$color <- v_color[Matrix::invPerm(
  p = order(
    x = grades$G3
  )
)]
```

```{r}
par(mfrow=c(2,2))
plot(
  x = lm_initial,
  pch = 19,
  col = grades$color
)

par(mfrow = c(2,2))
plot(
  x = lm_initialsquared,
  pch = 19,
  col = grades$color
)
```



Installing the fast dummies package to handle the ludacrous amount of dummy variables in the dataset

```{r}
library('fastDummies')
```

Creating the dummy variables
```{r}
gradedummy <- dummy_cols(grades, select_columns = c(
  'school', 'sex','address','famsize','Medu','Fedu','Mjob','Fjob','traveltime','studytime','schoolsup','famsup','activities','nursery','higher','internet','romantic','famrel','freetime','goout','Dalc','Walc','health'))
```

Building the linear Model with dummy variables
```{r}
lm_full <- lm(
  formula = G3^2 ~ G1 + G2 + age + failures + absences + school_GP + school_MS + sex_F + sex_M + address_R + address_U
  + famsize_GT3 + famsize_LE3 + Medu_0 + Medu_0 + Medu_1 + Medu_2 +Medu_3 + Medu_4 +Fedu_0 + Fedu_1 + Fedu_2 + Fedu_3 + Fedu_4 + Mjob_at_home + Mjob_health + Mjob_other + Mjob_services + Mjob_teacher + Fjob_at_home + Fjob_health + Fjob_other + Fjob_services + Fjob_teacher + traveltime_1 + traveltime_2 + traveltime_3 + traveltime_4 + studytime_1 + studytime_2 + studytime_3 + studytime_4 + schoolsup_no + schoolsup_yes + famsup_no + famsup_yes + activities_no + activities_yes + nursery_no + nursery_yes + higher_no + higher_yes + internet_no + internet_yes + romantic_no + romantic_yes + famrel_1 + famrel_2 + famrel_3 + famrel_4 + famrel_5 + freetime_1 + freetime_2 + freetime_3 + freetime_4 + freetime_5 + goout_1 + goout_2 + goout_3 + goout_4 + goout_5 + Dalc_1 + Dalc_2 + Dalc_3 + Dalc_4 + Dalc_5 + Walc_1 + Walc_2 + Walc_3 + Walc_4 + Walc_5 + health_1 + health_2 + health_3 + health_4 + health_5,
  data = gradedummy
)

summary(lm_full)
```

Diagnostic Plots
```{r}
v_color1 <- viridis::viridis(
  n = nrow(
    x = gradedummy
  )
)
gradedummy$color <- v_color[Matrix::invPerm(
  p = order(
    x = gradedummy$G3
  )
)]
```

```{r}
par(mfrow = c(2,2))
plot(
  x = lm_initialsquared,
  pch = 19,
  col = grades$color
)

par(mfrow=c(2,2))
plot(
  x = lm_full,
  pch = 19,
  col = gradedummy$color
)
```

Stepwise Regression

```{r}
library(olsrr)
```

Stepwise with p values
```{r}
stepwise.p <- ols_step_both_p(lm_full,prem = 0.05, details = FALSE)
stepwise.p
```

Stepwise with AIC values #Comment out when Knitting
```{r}
##stepwise.aic <- ols_step_both_aic(lm_full, details = FALSE)
##stepwise.aic
```

Reduced model post Stepwise p-value
```{r}
lm_reduced <- lm(
  formula = G3^2 ~ G2 + G1 + famrel_5 + age + studytime_1 + Fjob_services + Walc_1 + Medu_2,
  data = gradedummy
)

summary(lm_reduced)
```

```{r}
lm_reduced_FINAL <- lm(
  formula = G3^2 ~ G2 + G1 + famrel_1 + famrel_2 + famrel_3 + famrel_4 + famrel_5 + age + studytime_1 + studytime_2 + studytime_3 + studytime_4  + Fjob_services + Fjob_at_home + Fjob_health + Fjob_other + Fjob_teacher  + Walc_1 +  Walc_2 + Walc_3 + Walc_4 + Walc_5 + Medu_1 + Medu_2 + Medu_3 + Medu_4,
  data = gradedummy
)

summary(lm_reduced_FINAL)
```

Diagnostic Plots for Final Model
```{r}
par(mfrow=c(2,2))
plot(
  x = lm_reduced_FINAL,
  pch = 19,
  col = gradedummy$color
)
```

Comparing the models
```{r}
anova(lm_reduced,lm_reduced_FINAL)

anova(lm_initialsquared,lm_reduced_FINAL)
```

Predicted vs Observed
```{r}
pred <- predict(lm_reduced_FINAL)

dt_prediction = (data.frame((sqrt(pred)),(gradedummy$G3)))
head(dt_prediction,30)

plot(x=sqrt(pred),y=gradedummy$G3,xlab='Predicted Values',ylab='Actual Values',main='Predicted vs Observed')
abline(
  reg = lsfit(
    x = sqrt(pred),
    y = gradedummy$G3
  )
)
```






